---
title: "Persistent JavaScript References"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Persistent JavaScript References}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(QuickJSR)
```

QuickJSR normally converts JavaScript results to R objects immediately. This
is convenient for scalar results and final outputs, but repeated conversion is
expensive when large values pass through several JavaScript operations.

A `JSValueRef` keeps a value inside its JavaScript context. Property access and
function calls can then operate on the original value without converting it to
R. Conversion only occurs when `js_ref_to_r()` is called.

## Context profiles

`JSContext$new()` uses the `"host"` profile by default. It includes the
QuickJS `std` and `os` modules, console helpers, and the global `R` bridge.
Use `"standard"` when JavaScript needs the standard modules but not the global
R bridge, or `"bare"` for only the core JavaScript runtime.

```{r}
bare_ctx <- JSContext$new(profile = "bare")
bare_ctx$source(code = "function add(a, b) { return a + b; }")
bare_ctx$call("add", 2, 3)
```

All profiles support ordinary R-to-JavaScript value transfer. References
remain owned by the context that created them.

## Persistent and compiled evaluation

Use `ctx$eval()` when source should execute in an existing context and return
an R value. Unlike `qjs_eval()`, it preserves global state between calls.

```{r}
bare_ctx$eval("globalThis.count = 0; count")
bare_ctx$eval("count += 1; count")
```

Use `ctx$compile()` when the same global script will execute repeatedly.
Compilation parses the source without running it. `js_script_run()` converts
each result to R, while `js_script_run_ref()` retains the result in JavaScript.

```{r}
increment <- bare_ctx$compile("count += 1; count")
js_script_run(increment)
js_script_run(increment)

array_script <- bare_ctx$compile("[count, count + 1]")
array_result <- js_script_run_ref(array_script)
js_ref_to_r(array_result)
```

Compiled scripts take no arguments. For parameterised repeated work, evaluate
a function once with `eval_ref()` and invoke it with `js_ref_call()`.

```{r}
double <- bare_ctx$eval_ref("value => value * 2")
js_ref_to_r(js_ref_call(double, 21))
```

## Creating references

A context can create references by evaluating code, retrieving a global value,
or calling a function.

```{r}
ctx <- JSContext$new()
ctx$source(code = paste(
  "function identity(x) { return x; }",
  "function doubleValues(x) { return x.map(value => value * 2); }",
  "function sumValues(x) { return x.reduce((a, b) => a + b, 0); }"
))

evaluated <- ctx$eval_ref("[1, 2, 3, 4]")
retrieved <- ctx$get_ref("Math")
called <- ctx$call_ref("doubleValues", evaluated)

evaluated
called
```

The three context methods are:

| Method | Operation |
|---|---|
| `ctx$eval_ref(code)` | Evaluate code and retain its result |
| `ctx$get_ref(name)` | Retrieve a global value or dotted property path |
| `ctx$call_ref(name, ...)` | Call a global function or dotted method and retain its result |

## Passing references back to JavaScript

A reference from a context can be passed to ordinary or reference-returning
calls on that same context. The referenced value is reused directly.

```{r}
ctx$call("sumValues", evaluated)

doubled <- ctx$call_ref("doubleValues", evaluated)
ctx$call("sumValues", doubled)
```

This is the main performance benefit: intermediate arrays and objects do not
cross the R and JavaScript boundary.

## Property access and callable references

Use `js_ref_get()` to retain a property and `js_ref_call()` to call a retained
function. Both return another `JSValueRef`.

```{r}
pipeline <- ctx$eval_ref(
  "({
    values: [1, 2, 3],
    scale(multiplier) {
      this.values = this.values.map(value => value * multiplier);
      return this;
    },
    total() {
      return this.values.reduce((a, b) => a + b, 0);
    }
  })"
)

scale <- js_ref_get(pipeline, "scale")
scaled <- js_ref_call(scale, 10)
total <- js_ref_call(js_ref_get(scaled, "total"))

js_ref_to_r(total)
```

The reference returned for `scale` retains `pipeline` as its method receiver,
so JavaScript evaluates the call with the correct `this` value. Dotted context
paths retain receivers in the same way.

```{r}
ctx$source(code = "globalThis.counter = { value: 5, read() { return this.value; } }")
read <- ctx$get_ref("counter.read")
js_ref_to_r(js_ref_call(read))
```

## Materialising a value in R

Call `js_ref_to_r()` when the value is needed in R.

```{r}
js_ref_to_r(doubled)
```

Materialisation uses the normal QuickJSR conversion rules. Keeping it explicit
makes the location and cost of conversion visible.

## Context ownership

Every reference belongs to one context. Passing it to another context raises
an error because QuickJS values cannot move between runtimes.

```{r error = TRUE}
other_ctx <- JSContext$new()
other_ctx$source(code = "function identity(x) { return x; }")
other_ctx$call("identity", evaluated)
```

A reference keeps its original context alive. It remains valid even if the
context object is no longer otherwise reachable from R.

```{r}
detached <- local({
  local_ctx <- JSContext$new()
  local_ctx$eval_ref("({ answer: 42 })")
})

gc()
js_ref_to_r(detached)
```

## Cyclic and non-convertible values

References can retain cyclic objects, functions, promises, maps, sets, and
class instances without flattening them into R data structures.

```{r}
cycle <- ctx$eval_ref(
  "(() => {
    const value = { name: 'cycle' };
    value.self = value;
    return value;
  })()"
)

self <- js_ref_get(cycle, "self")
self
```

Do not materialise a cyclic graph unless the conversion policy supports its
shape. Continue operating on it through references instead.

## Choosing between values and references

Use ordinary `call()` and `get()` for scalars, small final results, and values
that are immediately needed in R. Use references when values are large, when
several JavaScript operations are chained, or when JavaScript identity and
structure must be preserved.

A typical high-performance workflow is:

1. Transfer or create data once with `eval_ref()`, `get_ref()`, or `call_ref()`.
2. Chain JavaScript work with `js_ref_get()` and `js_ref_call()`.
3. Return small scalar summaries through ordinary `call()` where practical.
4. Materialise large results once with `js_ref_to_r()`.
